---
title: "ESM 206 Lab 5"
author: "Tommy King"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---
Code Folding allows there to be an option for hiding or showing code in the knitted document.
```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Confidence Intervals

In this section, we will learn to use the 't.test()' function as a away to return the confidence interval using the t-distribution. 

## Example 1

Create a mock sample of observation values, and find the confidence interval. Then, we'll learn how to report that.

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 102)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)

```

### Risky & Tedious Way to present relevant statistics for pika mass: 

Mean pika mass at Paiute Pass is 91.9g $\pm$ 13.5g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g. 

### A high initial investment, but safer and better way to present relevant statistics:

In-line code referencing to stored objects! 

Use a single backtick on either side of a lowercase r to create an in-line bit of code. 

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)` g $\pm$ `r round(pika_sd,1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r pika_ci$conf.int[1]`]

Note: Use function my_list$name to return just the name of the list within the console
ALso for indexing (meaning pulling separate pieces of data from the inputted data and see what it returns), can use the function pika_mass[x] to show value at position x. 


# Two -sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0). 

- H0: Means difference = 0
- HA: Means difference is NOTE = 0

This is a two sided t-test because no directionality is implied. 

## Example 1 

```{r}
oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7 )

pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)

```

Is there a significant difference in mean heights for oak and pine trees?

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

The p-value of `r trees_t$p.value` means that there is a 1.8% chance of finding sample means *at least as different* as those I found by random chance if they were drawn from populations with the same mean height.

## Example 2

Using `mpg` dataset to compare city gas mileage for SUV's and compact cars. 

```{r}
# create a subset called compact_suv that only contains observations where the class is suv or compact

compact_suv <- mpg %>% 
  filter(class %in% c("suv", "compact"))
```

```{r}
ggplot(data = compact_suv, aes(x = cty)) + 
  geom_histogram(aes(fill = class), bins = 12) + 
  facet_wrap(~ class)

# creating a QQ plot
ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq(aes(color = class)) + 
  facet_wrap(~ class)
```
```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(mean_city = mean(cty), 
            sd_city = sd(cty),
            sample_size = n())

car_stats
```
```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty) # this will tell the code which values to pull from

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)
```

stats speak: reject the null hypothesis of equal city gas mileages

Mean gas Mileage for compact cars and SUV's differs significantly. (t(`r cars_t$parameter`) = `r cars_t$statistic`, p < 0.001, $\alpha$ = 0.05)
